---
title: "chapter5"
author: "Tiina Autio"
date: "2/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Analysis of human data

This analysis focuses on health, knowledge and empowerment indicators from various countries.  

```{r, echo=FALSE}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", header = TRUE, sep = ",")

str(human)
glimpse(human)
```
Data set contains 155 observations and 8 variables:  
- "GNI" = Gross National Income per capita  
- "Life.Exp" = Life expectancy at birth  
- "Edu.Exp" = Expected years of schooling   
- "Mat.Mor" = Maternal mortality ratio  
- "Ado.Birth" = Adolescent birth rate  
- "Parli.F" = Percetange of female representatives in parliament  
- "Edu2.F" = Proportion of females with at least secondary education  
- "Edu2.M" = Proportion of males with at least secondary education  
- "Labo.F" = Proportion of females in the labour force  
- "Labo.M" " Proportion of males in the labour force  
- "Edu2.FM" = Edu2.F / Edu2.M  
- "Labo.FM" = Labo2.F / Labo2.M  

Original data can be found here: http://hdr.undp.org/en/content/human-development-index-hdi  

Summary of the data and some visualisations:
```{r, echo=FALSE}
summary(human)
pairs(human)

library(GGally)
ggpairs(human)

cor(human) %>% corrplot()
```

## Principal component analysis (PCA)

First let's perform principal component analysis (PCA) to non-standardized data set. 
```{r, echo=FALSE}
pca_human <- prcomp(human)

s <- summary(pca_human)
pca_pr <- round(100*s$importance[2,], digits = 1)
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")

biplot(pca_human, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])
```

Then let's do the PCA again to standardized data set.
```{r, echo=FALSE}
human_std <- scale(human)

summary(human_std)

pca_human_std <- prcomp(human_std)

s <- summary(pca_human_std)
pca_pr <- round(100*s$importance[2,], digits = 1)
pc_lab <- paste0(names(pca_human_std), " (", pca_pr, "%)")

biplot(pca_human_std, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])
```
