---
title: "chapter4"
author: "Tiina Autio"
date: "2/16/2017"
output: html_document
---

# Analysis of Housing Values in Suburbs of Boston

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r , echo=FALSE}
library(MASS)
data("Boston")
```

Structure and summary of data:
```{r , echo=FALSE}
str(Boston)
summary(Boston)
```

Boston data set has 506 observtions and 15 variables. 

Columns:  

- crim: per capita crime rate by town.  

- zn: proportion of residential land zoned for lots over 25,000 sq.ft.  

- indus: proportion of non-retail business acres per town.  

- chas: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise).  

- nox: nitrogen oxides concentration (parts per 10 million).  

- rm: average number of rooms per dwelling.  

- age: proportion of owner-occupied units built prior to 1940.  

- dis: weighted mean of distances to five Boston employment centres.  

- rad: index of accessibility to radial highways.  

- tax: full-value property-tax rate per \$10,000.  

- ptratio: pupil-teacher ratio by town.  

- black: 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town.  

- lstat: lower status of the population (percent).  

- medv: median value of owner-occupied homes in \$1000s.  

(Source: https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/Boston.html)

Very rough visualisation of data distributions:
```{r , echo=FALSE}
pairs(Boston)
```
  
  
Visualiasation of data correlations:
```{r , echo=FALSE}
library(dplyr)
library(corrplot)
cor_matrix<-cor(Boston) 

# print the correlation matrix
cor_matrix %>% round(2)

# visualize the correlation matrix
corrplot(cor_matrix, method="circle", type="upper")
```

Seems that crime correlates with rad, tax, lstat, nox, indus, age and ptratio. Crime correlates least with black and medv (-0.39) and almost as much with distance (-0.38).  

## Standardizing data

Summary of standardized data set:
```{r , echo=FALSE}
scaled <- scale(Boston)
summary(scaled)
cor(scaled)
```
Variables have change so that variables with strongest crime correlations have even more stronger correlation. Weakest correlations haven't change so much.  

Let's create categorical variable of crime rate. First we create quantile vector of crime and give following categories for it: "low", "med_low", "med_high", "high". After that I'll remove the original crime rate variable from the data set. Also I divide the data set to train and test sets.

```{r , echo=FALSE}

scaled <- as.data.frame(scaled)
class(scaled)
str(scaled)
scaled_crim <- scale(scaled$crim)

q <- quantile(scaled_crim)

# create a categorical variable 'crime'
crime <- cut(scaled_crim, breaks = q, include.lowest = TRUE, label = c("low", "med_low", "med_high", "high") )
table(crime)

# remove original crim from the dataset
scaled <- dplyr::select(scaled, -crim)
scaled <- data.frame(scaled, crime)

n <- nrow(scaled)
sample <- sample(n, size = n * 0.8)

train <- scaled[sample,]
test <- scaled[-sample,]

correct_classes <- test$crime

test <- dplyr::select(test, -crime)
```

Then I'll make the linear discriminant analysis on train set using categorical crime rate as target variable. This is the LDA (bi)plot:
```{r , echo=FALSE}

lda.fit <- lda(crime ~ ., data = train)

# the function for lda biplot arrows
lda.arrows <- function(x, myscale = 1, arrow_heads = 0.1, color = "red", tex = 0.75, choices = c(1,2)){
  heads <- coef(x)
  arrows(x0 = 0, y0 = 0, 
         x1 = myscale * heads[,choices[1]], 
         y1 = myscale * heads[,choices[2]], col=color, length = arrow_heads)
  text(myscale * heads[,choices], labels = row.names(heads), 
       cex = tex, col=color, pos=3)
}

# target classes as numeric
classes <- as.numeric(train$crime)


# plot the lda results
plot(lda.fit, col = classes, pch= classes, dimen = 2)
lda.arrows(lda.fit, myscale = 1)

```

Now I remove crime categories from test and also the categorical crime variable from test.

```{r , echo=FALSE}

lda.pred <- predict(lda.fit, newdata = test)

# cross tabulate the results
table(correct = correct_classes, predicted = lda.pred$class)
```
