---
title: "chapter3"
author: "Tiina Autio"
date: "2/10/2017"
output: html_document
---

# Analysis of Student Alcohol consumption

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's take a look of the data:
```{r echo=FALSE}
alc <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt", sep=",", header=TRUE)
```

```{r echo=TRUE}
dim(alc)
colnames(alc)
str(alc)
summary(alc) 
```

The data consists of 382 observations and 35 variables.

Overall visualisation of the data:
```{r echo=FALSE}
library(tidyr); library(dplyr); library(ggplot2)
gather(alc) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()

```



## Hypothesis

From this data set I chose variables:  
- absenses  
- health  
- failures  
- reason

1) Hypothesis of absences:
H0: High absences are related to high alcohol consumption
H1: More absences aren't related to high alcohol consumption

2) Hypothesis of health:
H0: Poor health is related to high alcohol consumption
H1: Poor health isn't related to high alcohol consumption

3) Hypothesis of failures:
H0: More failures are related to high alcohol consumption
H1: Failures aren't related to high alcohol consumption

4) Hypothesis of activities:
H0: No activities is related to high alcohol consumption
H1: Activities are related to high alcohol consumption

Absenses:
I divided absences to logical type variable absences_high that is true if student have more than three absences (3.00 is the median value for absences).
```{r echo=FALSE}

alc <- mutate(alc, absences_high = absences > 3)
alc %>% group_by(absences_high, high_use) %>% summarise(count = n(), mean_grade = mean(G3))

g1 <- ggplot(alc, aes(x = high_use, y = G3, col = absences_high))

# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("absences_high")
```

Health:
```{r echo=FALSE}

alc <- mutate(alc, health_poor = health < 3)
alc %>% group_by(health_poor, high_use) %>% summarise(count = n(), mean_grade = mean(G3))

g1 <- ggplot(alc, aes(x = high_use, y = G3, col = health_poor))

# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("health_poor")
```


Failures:
```{r echo=FALSE}

alc <- mutate(alc, failures_high = failures > 1)
alc %>% group_by(failures_high, high_use) %>% summarise(count = n(), mean_grade = mean(G3))

g1 <- ggplot(alc, aes(x = high_use, y = G3, col = failures_high))

# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("failures_high")
```

Activities:
```{r echo=FALSE}

alc %>% group_by(activities, high_use) %>% summarise(count = n(), mean_grade = mean(G3))

g1 <- ggplot(alc, aes(x = high_use, y = G3, col = activities))

# define the plot as a boxplot and draw it
g1 + geom_boxplot() + ylab("activities")
```


```{r echo=FALSE}
m <- glm(high_use ~ failures_high + absences_high + activities + health_poor, data = alc, family = "binomial")

# predict() the probability of high_use
probabilities <- predict(m, type = "response")

# add the predicted probabilities to 'alc'
alc <- mutate(alc, probability = probabilities)

# use the probabilities to make a prediction of high_use
alc <- mutate(alc, prediction = probability > 0.5)

# tabulate the target variable versus the predictions
table(high_use = alc$high_use, prediction = alc$prediction)

loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

# call loss_func to compute the average number of wrong predictions in the (training) data
loss_func(class = alc$high_use, prob = alc$probability)
```

```{r echo=FALSE}
# define a loss function (average prediction error)
loss_func <- function(class, prob) {
  n_wrong <- abs(class - prob) > 0.5
  mean(n_wrong)
}

# compute the average number of wrong predictions in the (training) data
mean(loss_func(alc$high_use, alc$prediction))

# K-fold cross-validation
library(boot)
cv <- cv.glm(data = alc, cost = loss_func, glmfit = m, K = 10)
?cv.glm
# average number of wrong predictions in the cross validation
cv$delta[1]
```