---
title: "chapter5"
author: "Tiina Autio"
date: "2/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Analysis of human data

This analysis focuses on health, knowledge and empowerment indicators from various countries.  

```{r, echo=FALSE}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", header = TRUE, sep = ",")

str(human)
```
Data set contains 155 observations and 8 variables:  
- "GNI" = Gross National Income per capita  
- "Life.Exp" = Life expectancy at birth  
- "Edu.Exp" = Expected years of schooling   
- "Mat.Mor" = Maternal mortality ratio  
- "Ado.Birth" = Adolescent birth rate  
- "Parli.F" = Percetange of female representatives in parliament  
- "Edu2.F" = Proportion of females with at least secondary education  
- "Edu2.M" = Proportion of males with at least secondary education  
- "Labo.F" = Proportion of females in the labour force  
- "Labo.M" " Proportion of males in the labour force  
- "Edu2.FM" = Edu2.F / Edu2.M  
- "Labo.FM" = Labo2.F / Labo2.M  

Original data can be found here: http://hdr.undp.org/en/content/human-development-index-hdi  

Summary of the data and some visualisations:
```{r echo=FALSE}
summary(human)
pairs(human)

library(GGally)
library(dplyr)
library(corrplot)

ggpairs(human)

cor(human) %>% corrplot()
```

## Principal component analysis (PCA)

First let's perform principal component analysis (PCA) to non-standardized data set. 
```{r, echo=FALSE}
pca_human <- prcomp(human)

s <- summary(pca_human)
pca_pr <- round(100*s$importance[2,], digits = 1)
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")

biplot(pca_human, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])
```
Without scaling all the result are shown in the same area so that it is impossible to see the differences between countries. 


So let's standardized the data setthe and do the PCA again.
```{r, echo=FALSE}
human_std <- scale(human)

summary(human_std)
pca_human_std <- prcomp(human_std)

s <- summary(pca_human_std)
pca_pr <- round(100*s$importance[2,], digits = 1)
pc_lab <- paste0(names(pca_human_std), " (", pca_pr, "%)")

biplot(pca_human_std, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])
```
After scaling the PCA seems a lot more balanced and results are more easily to read. I interpret that the First two variables are Ado.Birth and Mat.Mor.  

## Tea data set

```{r, echo=FALSE}
library(FactoMineR)
data(tea)

str(tea)
summary(tea)

keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- select(tea, one_of(keep_columns))

library(ggplot2)
library(tidyr)
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

mca <- MCA(tea_time, graph = TRUE)
summary(mca)


```